---
title: "p8105_hw2_jz3573"
output:
 github_document 
date: "2023-10-03"
---


## Load library 
```{r setup}
library(tidyverse)
library(readxl)
library(tidyr)
```


# Quesioin 2

## tidy up Mr.Trash Wheel
```{r tidy up Mr.Trash Wheel}
Mr=
  read_excel("./data/Trash.xlsx", 1,range="A2:N586") %>% 
  janitor::clean_names() %>% 
  rename(dumpster_number=dumpster) %>% 
  drop_na(dumpster_number) %>% 
  mutate(homes_powered = weight_tons *500/30,
         trash_wheel="Mr")
 
skimr::skim(Mr)
```
## Tidy up Professor Trash Wheel
```{r tidy up Professor Trash Wheel}
Pro=
  read_excel("./data/Trash.xlsx", 2,range="A2:M108") %>% 
  janitor::clean_names() %>% 
  rename(dumpster_number=dumpster) %>% 
  drop_na(dumpster_number) %>% 
  mutate(homes_powered = weight_tons *500/30,
         trash_wheel="Professor")
```

## Tidy up Gwynnda
```{r tidy up Gwynnda Trash Wheel}
Gwy=
  read_excel("./data/Trash.xlsx", 3 ,range="A2:L157") %>% 
  janitor::clean_names() %>% 
  rename(dumpster_number=dumpster) %>% 
  drop_na(dumpster_number) %>% 
  mutate(homes_powered= weight_tons *500/30,
          trash_wheel="Gwynnda") 
  

 skimr::skim(Gwy)

```
## Join the three datasets
```{r join datasets}
joint=
     full_join(Gwy, Pro) |>
     full_join(x = _, y = Mr)
 
  
```

## Descriptiion of the dataset:
 
The dataset contains `r nrow(joint)` observations and `r ncol(joint)` variables.

The total weight of trash collected by Professor Trash Wheel is `r subset(joint, trash_wheel == "Professor") |> pull(weight_tons) |> sum()`

The dataset describe the detail of drumpsters, which include the drumpsters' number, the weight of trash and the type of trash. 

The total number of cigarette butts collected by Gwynnda in July of 2021 is `r `subset(joint, trash_wheel=="Gwynnda" &  month=="July" & year=="2021") %>% pull (cigarette_butts) %>% sum()`



# Question 3


