---
title: "p8105_hw2_jz3573"
output:
 github_document
date: "2023-10-03"
---


## Load library 
```{r setup}
library(tidyverse)
library(tidyr)
library(dplyr)
library(readxl)
```


# Quesioin 2

## tidy up Mr.Trash Wheel
```{r tidy up Mr.Trash Wheel}
Mr=
  read_excel("./data/Trash.xlsx", 1,range="A2:N586") %>% 
  janitor::clean_names() %>% 
  rename(dumpster_number=dumpster) %>% 
  drop_na(dumpster_number) %>% 
  mutate(homes_powered = weight_tons *500/30,
         trash_wheel="Mr")
 
```
## Tidy up Professor Trash Wheel
```{r tidy up Professor Trash Wheel}
Pro=
  read_excel("./data/Trash.xlsx", 2,range="A2:M108") %>% 
  janitor::clean_names() %>% 
  rename(dumpster_number=dumpster) %>% 
  drop_na(dumpster_number) %>% 
  mutate(homes_powered = weight_tons *500/30,
         trash_wheel="Professor")
```

## Tidy up Gwynnda
```{r tidy up Gwynnda Trash Wheel}
Gwy=
  read_excel("./data/Trash.xlsx", 3 ,range="A2:L157") %>% 
  janitor::clean_names() %>% 
  rename(dumpster_number=dumpster) %>% 
  drop_na(dumpster_number) %>% 
  mutate(homes_powered= weight_tons *500/30,
          trash_wheel="Gwynnda") 


```
## Merge the three datasets
```{r merge datasets}
joint=
     full_join(Gwy, Pro) |>
     full_join(x = _, y = Mr)
 
```

## Descriptiion of the dataset:
 
The dataset contains `r nrow(joint)` observations and `r ncol(joint)` variables.

The total weight of trash collected by Professor Trash Wheel is `r subset(joint, trash_wheel == "Professor") |> pull(weight_tons) |> sum()`

The dataset describe the detail of drumpsters, which include the drumpsters' number, the weight of trash and the type of trash. 

The total number of cigarette butts collected by Gwynnda in July of 2021 is `r subset(joint, trash_wheel=="Gwynnda" &  month=="July" & year=="2021") %>% pull (cigarette_butts) %>% sum()`



# Question 3

## Tidy baseline df
Following codes ensure that sex and APOE4 carrier status are appropriate encoded and every participant meet the stated inclusion criteria (i.e. no MCI at baseline).
```{r tidy baseline df }
baseline=
  read.csv("./data/data_mci/MCI_baseline.csv", skip = 1) %>% 
  janitor::clean_names() %>% 
  mutate(
      sex =case_match(sex,
                      0 ~ "female",
                      1 ~ "male"),
      apoe4=case_match(apoe4,
                      1 ~ "carrier",
                      0 ~ "non-carrier")) %>% 
     filter( age_at_onset=="." | age_at_onset > current_age)

```

## Description of baseline df
```{r}
x=row(baseline)
y=filter(baseline, age_at_onset != ".") |> nrow ()
z=pull(baseline, current_age) |>  mean()
a=scales::percent(nrow(filter(baseline, sex == "female" & apoe4 =="carrier")) / nrow(filter(baseline, sex == "female")))
```

479 participants were recruited
93 participants develop MCI? 
The average baseline age is 60.03
30% of women in the study are APOE4 carriers 



## Tidy up the longitudinal dataset:
Following codes tidy up the variables' names and transfer the dataframe into a machine-friendly frame. 
```{r tidy up amyloid}
amyloid=
  read.csv("./data/data_mci/mci_amyloid.csv", skip = 1) %>% 
  janitor::clean_names() %>% 
  rename ("id"="study_id") %>% 
  pivot_longer( 
    cols=baseline:time_8, 
    names_to = "time", 
    values_to = "amyloid_ratio") 
```

## Descripe the longitudinal dataframe:

The dataset contains `r nrow(amyloid)` observations and `r ncol(amyloid)` variables. There are  `r n_distinct((pull(amyloid,id)))` participants in this dataset. 
MCI is recorded at baseline, time 2, time 4, time 6, and time 8. There are a lot of missing values, meaning that many participants missed visits. 

## Merge the datasets and its descriptions
The two datasets were joined by the common variable: id. Only those participants who are enrolled in both datasets are shown in the full_df. 
```{r merge the datasets}
full_df = 
  inner_join(baseline, amyloid, by = c("id"))
```

The dataset contains `r nrow(full_df)` observations and `r ncol(full_df)` variables.There are `r n_distinct(pull(full_df, id))` participants in the dataframe. 

`r nrow(baseline)-n_distinct(pull(full_df, id))`participants appear in only the baseline datasets. 
 `r n_distinct(pull(amyloid, id))-n_distinct(pull(full_df, id))`participants appear in only the amyloid datasets.
 

## Export the dataset as a csv
```{r}
write.csv(full_df, "./data/baseline_amyloid_merged.csv")
```
```


